SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations [v17]

### 2015-09-03 paper.tex
Anca's comments

\author{Michael Laskey$^1$, Sam Staszak$^3$, Wesley Yu-Shu Hsieh$^1$, Jeff Mahler$^1$, Florian T. Pokorny$^1$, Anca Dragan$^1$, Ken Goldberg$^{1,2}$% <-this % stops a space
\thanks{$^1$ Department of Electrical Engineering and Computer Sciences; {\small \{mdlaskey,iamwesleyhsieh, ftpokorny, anca\}@berkeley.edu}}%
\thanks{$^2$ Department of Industrial Engineering and Operations Research; {\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley;  Berkeley, CA 94720, USA}%
}
\begin{document}

@@

\begin{abstract}
Online algorithms for robot model-free learning from demonstration (such as DAgger) can learn policies for problems where system
dynamics and cost function are unknown. However, during learning they impose a burden on supervisors to provide labels (control
signals) as the robot encounters new states when iteratively executing its current best policy. We introduce the SHIV
algorithm (Svm-based reduction in Human InterVention), which reduces this burden by enabling the robot to only request supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary high dimensional state distributions.  To facilitate scaling and outlier rejection, risk is based on distance to an approximate level set boundary defined by a One Class support vector machine.  We compare SHIV with DAgger in three policy learning experiments: 1) a driving simulator with a 27936D visual HOG feature state representation, 2) 4DOF robot arm push-grasping in clutter (simulation with Box2D) with a 22D state representation  of joint angles and the pose of each object, and 3) surgical needle insertion with a 16D state represented spatial transform.  Results suggest that SHIV can reduce the number of queries up to 70$\%$.

@@

In model-free robot learning from demonstration, a robot learns to perform a task, such as driving or grasping an object in a cluttered environment as shown in Fig. \ref{fig:teaser}, from examples provided by a competent supervisor, usually a human. The problem is model-free if  the system dynamics are unknown, eg. dynamics model may be highly non-linear or noisy. In learning from demonstration, the cost function, which is the negative of the reward function, is usually unknown or significantly delayed. In the grasping for clutter example, shown in Fig. \ref{fig:teaser}(b) the robot is only penalized when an object is knocked off the table, which means a potentially large number of states can be visited before any feedback. To potentially avoid this a supervisor provides guidance through demonstrations and the robot learns a policy mapping states to controls~\cite{argall2009survey}.

In the supervised setting, a policy is learned offline from a fixed set of examples. When the robot executes this policy, the sequential process can quickly evolve away from the previously explored set of states.  The learned policy might not be helpful in unfamiliar states, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will put the robot into states near the side of the road where it has not encountered before, where its policy could  fail~\cite{pomerleau1989alvinn}. It has been shown, the number of errors made by the robot in such cases  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/teaser.pdf}
\caption{ 
\todo{update figure}
SHIV reduces the number of queries to the supervisor in our three test domains: 1) a driving simulator where the goal is to learn a controller on HOG features extracted from synthetic images to keep a car on a polygonal track; 2) push-grasping in clutter in a physics simulator, where the goal is to learn a controller from human demonstrations to grasp the green object while not letting the red square objects fall off the gray boundary representing a table; 3) learning a controller for the Da Vinci Research Kit from a master surgeon's demonstrations of correcting a needle before suturing in surgery. 
}
\vspace*{-10pt}
\label{fig:teaser}
\end{figure}

On-line learning from demonstration approaches such as DAgger iteratively gather more examples from the supervisor~\cite{grollman2007dogged,ross2010efficient,ross2010reduction}. In DAgger, a series of policies are learned.  In each iteration, the latest policy is "rolled out" (executed) and the supervisor provides demonstrations for all  new states, which are then aggregated with the previous examples for the next iteration. DAgger and related algorithms have been applied in a wide range of applications, from quadrotor flight to natural language to Atari games~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}. It has been shown for DAgger, the number of errors to scale linearly for strongly convex loss functions.

One drawback is that such online algorithms increase the burden on the supervisor. Our goal is to reduce supervisor burden without degrading robot performance, by only requesting demonstrations for risky states.

Risk arises because:  (1) states are different than the states previously encountered and used to generate the current policy, i.e., they can be considered novel or outliers~\cite{hodge2004survey}; or (2) states are in a region that have been misclassified, in previous training data.
%\adnote{Ken's version below does not seem to represent what we actually do.}
%the action associated with a state under the current policy yield a resulting state that is inconsistent with the expected next state. [how do we measure 2? - Ken]

We introduce the SHIV algorithm (Svm-based reduction in Human InterVention), which can reduce supervisor burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary state distributions.. It builds on the One Class SVM method for approximating quantile level sets \cite{scholkopf2001estimating} to provide a classification method for deciding if a state is risky or safe.

% \adnote{Old intro:}

% Consider the problem of teaching a robot a complex task such cooking a meal or as racing a toy car around a track
% against human opponents. When the robot is only provided with visual demonstrations without the knowledge or an
% underlying cost function describing the task, traditional model based control approaches are often not applicable. In
% these situations, the learning from demonstrations (LfD) approach provides a paradigmn for teaching a robot how to
% accomplish the task \cite{ross2013learning,pomerleau1989alvinn,schulman2013case}.\adnote{these are very random things to cite for LfD, FYI - there is a survey by Argall that usually gets cited}

% One approach to LfD is to train a policy, or a function estimator (either a classifier or regressor), to predict a
% supervisor's behavior given training data consisting of pairs of observations (input) and actions (output) performed
% while accomplishing a task. Since the robot's actions typically affect the resulting next state, the resulting sequences
% of states encountered during the execution of a policy are not independent random variables, thus making statistical
% inference challenging. Ross et al. proved that this can cause the error in the policy estimator to compound over the
% time and lead to performance degradation that is proportional to the time horizon of the task squared. The intuition
% behind this result is that as the robot makes mistakes with respect to the supervisor's policy it drifts from the
% distribution it was trained on and might encounter new states it can't generalize to.  Ross and Bagnell overcome this by
% presenting an iterative method that first trains a policy on the data observed using standard supervised learning
% techniques then tries (or 'rolls out') the policy out in the environment to see what states are likely to occur under
% the robot's trained policy.  The supervisor then tells the robot what controls it should applied after each iteration
% and the policy is retrained \cite{ross2010reduction}. The DAgger method has been successfully used since to teach a
% robot how to follow verbal instructions, achieve state of the art performance on Atari games and fly a quad copter
% through a forest~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}.

% During learning, implementations of DAgger require the supervisor to manually provide control inputs that the robot
% should have applied for each states it visits when rolling out the current best policy estimate. If the supervisor is a
% human or an algorithm with a high time complexity, this can be tedious and expensive. We are hence interested in
% reducing the burdon on the supervisor by partitioning the state space into states where the policy can either be
% executed autonomously with little expected risk, or where supervisor control input (or 'labels') are likely to be
% required due to high 'risk'. A state can be risky for 2 reasons: 1) we encounter a state which has a low density of
% training data in its vicinity, which potentially means our policy will mis-predict the supervisor
% \cite{tokdar2010importance} 2) the state is near previously visited states, but the control-input is mis-predicted
% nonetheless. If an accurate density estimation corresponding to
% the pairs of visited states and control signals is available, the first scenario can be captured by regions of
% state-space with a low density estimate. Areas of low density are then considered to be risky since an insufficient data
% is available to accurately model the demonstrator's policy. This idea is also captured in the notion that 'a trained
% model will be able to generalize within the distribution it is trained on' \cite{tokdar2010importance} in statistical
% machine learning. The second case arises for example when the current best policy estimate is locally inconsistent with supervisor
% demonstrations.
% %
% %Vague!!!
% %This property can be seen used in importance sampling when data is collected from one distribution, but tested on
% %another and has to be re-weighted %to the test distribution \cite{huang2006correcting}. 
% %
% While the amount of data needed to accurately estimate the density increases exponentially with the dimension of state
% space \cite{nadaraya1964estimating}, we can turn to approximate methods which instead attempt to learn an approximate
% representation of the quantile levelsets of the distribution. Thus, we propose using a modified version the One Class
% SVM method for support estimation that only tries to estimate the level set of the quantile function of the density
% \cite{scholkopf2001estimating} at a user-determined quantile threshold which has been used even with high-dimensional
% image data \cite{liu2014unsupervised}. Our modification furthermore removes areas of state space from the super-levelset
% estimate where the local policy estimate is inconsistent with the supervisor demonstrations.

% To apply our approach, we make the same assumptions as in the case of DAgger: 1) we require access to an environment
% where a robot can try out (roll out) policies to sample trajectories and where we are able to collect observation states
% and corresponding control inputs along these trajectories. 2) an interface allowing a supervisor to provide control
% inputs that should have been applied for states along parts of the rolled out trajectories. 3) We assume that we are
% able to collect a series of observations from the environment and the controls applied. 4) Lastly, for our theoretical
% analysis to apply, we assume the current observation only depends on the previous observation and control input (i.e.
% the Markovian assumption).

% Our core contribution is the extension of DAgger to \acro~which incorporates a notion of risk to learning from
% demonstrations and reduces supervisor
% burdon by asking for supervisor demonstrations only in `risky' states. 
% We achieve this by proposing a modified version of the One Class SVM, that both estimates the quantile level sets 
% of the distribution of states with demonstrations and additionally
% and carves out areas in the support where we know our policy will mis-predict the supervisor, thus identifying states of high
% risk.  Our experimental results on a Super Mario Atari video game,  suggest \acro~is able to reach DAgger's
% peak performance of surviving for an average of 2200 simulation steps with only 1200 queries to the supervisor, while DAgger required 2800 queries
% averaged over 20 levels.
% {\color{blue}[-- FINALIZE --]}

@@

\noindent\textbf{Active Learning from Demonstration}
Online learning from demonstration can be formulated as a stream-based active learning problem ~\cite{atlas1990training,cohn1994improving}. In stream based active learning, the decision of whether to query the supervisor (for a control signal or not) is not over the entire state space (like in traditional pool-based active learning), but only on the states encountered during learning.

@@

In SHIV, we use a query selection method that allows for a non-stationary distribution, evaluating risk based on novelty (is the state an outlier to the training distribution), as in Fig. \ref{fig:support_example}(c). Additionally, we also assign higher risk to  historically mislabeled regions, as in Fig. \ref{fig:support_example}(d). We then use a classification method based on the One Class SVM \cite{scholkopf2001estimating} to predict if a state is risky or not risky, which we we define as safe.%Our approach builds on the One Class SVM method for approximating quantile level sets \cite{}, which has already been used in novelty detection on high-dimensional image data \cite{}.

\noindent\textbf{Risk via Novelty Detection.}
One part of our risk definition is the notion that a trained model will be able to generalize within the distribution it is
trained on \cite{tokdar2010importance}. Novelty detection \cite{hodge2004survey} is concerned with recognizing that a sample is outside of this distribution.
One approach to novelty detection is to directly estimate the underlying probability density function from which the data is sampled. If the probability of that point is low with respect to the underlying density, it will be marked as novel. However, the amount of data needed to accurately do so scales exponentially in the number of dimensions \cite{nadaraya1964estimating}.

An alternative to evaluating the probability of a data point is to measure distance to its nearest neighbors \cite{knox1998algorithms}. However, this approach was shown to be
susceptible to outliers since nearest neighbors  incorporates only local information about the data. Thus, a group of outliers can be close together, but significantly far from the majority of the data and nearest neighbors would mark them as not novel \cite{hodge2004survey}.

Our approach is based on the One Class SVM proposed by Scholk{\"o}pf et al., which estimates a particular quantile level set for the training data by solving a convex quadratic program  to find support
vectors~\cite{scholkopf2001estimating}. The method has been theoretically shown to approximate the quantile levelset of a density estimate
asymptotically for correctly chosen bandwidth settings and in the case of a normalized Gaussian kernel function \cite{vert2006consistency}. 
In \cite{liu2014unsupervised}, the One Class SVM has furthermore been used for novelty detection in high-dimensional image data. To our knowledge the use of One Class SVM to online learning from demonstration has never been done before. 
\begin{figure*}[t]
\centering

@@

\caption{\todo{update figure with elipse} A comparison of different query selection strategies for active learning on a non-stationary distribution. The shaded and empty circles are training data from two classes, $0$ and $1$ respectively. The red empty circles are samples from a new distribution (produced by executing the learned policy), and belong to class $1$. Typical strategies assign high risk to states close the decision boundary (a), or for which a set of estimators disagree (b). Neither of these apply to our new samples in red. In contrast, we use a strategy that is amenable to non-stationary distributions by assigning high risk to novel states (c) and states in historically mislabeled regions. 
%Confidence and Query by Committee methods measure
%{\color{blue} FP:estimate} states far away from the decision boundary of the estimator as having low risk,
%however novelty detection methods measure {\color{blue} FP:estimate} the level of risk as
%{\color{blue}FP:by} how far a point is from the trained data. Our
%modified risk method attempts to measure risk both as a function of how far it
%{\color{blue}it gramatically refers to risk,
%but you mean sample "it$\mapsto$a sample"?} is from data trained on
%{\color{blue}grammatically strange$\mapsto$from the training data} and the decision boundary of the estimator. 
}
\label{fig:support_example}
\end{figure*}

% The field of learning from demonstrations has achieved state of the art performance in a variety of areas within robotics
% \cite{argall2009survey}. Abbeel et al.~used an iLQR controller based on helicopter trajectories demonstrated by a human supervisor
% to successfully learn a series of aerobatics stunts \cite{abbeel2007application}. Billard and Matari used a hierarchy of
% neural networks on video data and tracking markers to imitate the movement of human arms on a 37 degree of freedom humaniod
% robot \cite{billard2001learning}. Schulman et al. used a non-linear warping technique based on thin-plate splines
% to transer demonstrations of a human controlling a PR2 robot to tie knots with a rope to new and unseen initial
% configurations of the~\cite{schulman2013case}.

% A subset of research in the field of learning from demonstrations is working within the assumptions that the dynamics of
% the robot and the environment are not known and that no access to a cost function describing a given task is available.
% Under these assumptions, a set of trajectories and controls can be collected from a supervisor and the policy learning
% problem can be treated as a supervised learning problem that determines a function from state space to
% controls~\cite{argall2009survey} by means of regression. Pomerleau et al. used this approach on raw image data and
% associated control inputs provided by a person driving a car. The authors then learned a policy that allowed a car to
% travel on a highway. However, they observed that if the car deviated from the middle of the road the learned policy was
% not able to recover, this behavior was attributed to the fact the car only drove in middle of the road while collecting
% demonstrations~\cite{pomerleau1989alvinn}. Ross et al. studied this effect and showed that error can accumulate at a
% rate quadratic in the time horizon of the policy in general. Ross et al. then proposed a learning algorithm, SMILE, that
% stochastically mixes the supervisor's control with the robot's policy control signal during training, thus allowing the
% robot to explore states the supervisor had not visited during training~\cite{ross2010efficient}.

% Ross and Bagnell improved upon this method with DAgger (Dataset Aggregation) -- an iterative method that first trains a
% policy on the data observed using standard supervised learning techniques and then tries, or 'rolls out' the policy in
% the environment to see what states are likely to occur under the policy. The supervisor then tells the robot what
% controls it should applied at each stage of this roll-out and the policy is retrained with an aggregate of all data seen
% before \cite{ross2010reduction}. This algorithm has found widespread adoption in the robotics community. in
% \cite{ross2013learning}, a quad-copter using DAgger learned how to fly through a forest using only Histogram of Gradient
% (HOG) features extracted from camera images \cite{ross2013learning}. Guo et al. combined DAgger with Deep Q-learning
% methods to achieve state of the art performance on a common Atari game benchmark in Reinforcement Learning
% \cite{NIPS2014_5421}. Duvallet et al. furthermore used DAgger to teach robots to follow verbal instructions about where
% to move sematically in a building with a training set consisting of of humans following similar verbal instructions
% \cite{duvallet2013imitation}. Levine et al. very recently extended the iterative learning paradigm through Guided Policy
% Search, which replaces the supervisor with iLQG and uses KL-divergence constraints to train the policy~\cite{levine2015end}.

% One limitation of DAgger which we address in the present work is that it requires the demonstrator, or supervisor, to
% provide control inputs that the supervisor would have applied for every state encountered along a roll-out of a trajectory at every interation of the
% algorithm. Providing these control inputs can be very tedious or expensive both in the case where the `supervisor' is a
% human or an algorithm with a high computational demand.

% In prior work, Judah et al. proposed applying active learning techniques to improve the learninng ability of an algorithm given a fixed budget. 
% \cite{judah2011active,judah2012active}. A drawback of this approach is that it requires us to estimate a complete
% density of states from the samples, which can be difficult especially in high dimensions. Furthermore, the  
% at each iteration only one labeled state was given. Thus, the robot could potentially be required to try out many more iterations of learning in order to successfully accomplished the task.

% \begin{enumerate}
% \item Our work is different than traditional pool base active learning for two reasons 1) the data comes in as a stream (i.e. we decided whether to label or not label the sample vs. using a pool base methods) 2) the estimator being trained is part of the parametrization of the distribution of states being encountered at run time, thus violating the i.i.d assumption common in these methods. 
% \item Some techniques that can handle these approaches are 1) uncertainty sampling , or measuring some confidence level in a classifier and deciding whether to label or not. 2) Query by committee, which involves sampling a committee of hypothesises from a version space of consistent classifiers 3) Density weighted model, which fitting a density to the unlabeled pool and then using a technique such as uncertainty sampling or query by committee to chose the label with the highest uncertainity times the density under the model. The reason for density weighted model is because the previous approaches are prone to labeling outliers \cite{settles2008analysis}. 
% \item Our approach can be considered a query selective strategy that uses the ideas of density weighted models. The level set estimation of the One Class SVM is a binary decision metric of how close a point is to the global support points that define its neighbors. Then our hole punching technique places a "prior" on points that are likely to be misclassified condition on the fact that points nearby are misclassified. Thus, encapsulating both a notion of uncertainity and distance to relative points. 
% \item One point that I've been thinking about is that traditional active learning is about training an estimator on some fix distribution and then trying to predict how the estimator will change given new points sampled from the same distribution. Thus most approaches work on the estimator themselves (query by committee sampling different hypotheses or uncertainty sampling measuring confidence). This assumption is violated in DAgger because you are sampling from a new distribution each iteration different from what is trained on. Thus, a confidence threshold such as (distance to hyperplane for an svm) would be misleading because if the point was drawn from a different distribution it could be very far from the hyperplane, but have a different label than that side of the hyperplane. Thus, we purpose not looking at the estimator but at the data itself. We make an assumption that points close to other points in our feature space will have similar labels, but the outliers are unknown and will need new labels. I suggest we run a uncertainty sampling technique vs. SHIV to verify all this.

% \end{enumerate}

% %[Let's discuss this part]
% {\color{blue}[-- IMPROVE THIS PARAGRAPH --]} In order to determine state space regions where a policy is likely to determine
% an incorrect control input, we leverage a
% property in statistical machine learning that a trained model will be able to generalize within the distribution it is
% trained on \cite{tokdar2010importance}. This property is used in importance sampling when data is collected from one
% distribution, but tested on another and has to be re-weighted to the test distribution \cite{huang2006correcting}. One
% approach is to estimate the probability density of states the policy is trained on. However, the amount of data needed to accurately estimate the density scales exponentially in the number of \cite{nadaraya1964estimating}.

% Several alternative approaches have been proposed to determine regions of state-space where an estimator is likely fail
% to generalize away from the training-data \cite{markou2003novelty}. Knox and Ng \cite{knox1998algorithms} used a nearest
% neighbor-based approach by estimating if a point was close to $k$ neighbors, however this approach was shown to be
% susceptible to outliers since nearest neighbors mostly incorporate local information about the data. Manevitz and Yousef
% \cite{manevitz2002one} trained a neural network filter to try and reconstruct the data and when the reconstruction error
% at a point was high, the point was considered as likely to be outside of the estimators domain within which
% generalization was possible. Manevitz and Yousef conjecture  that finding the right architecture can be task specific
% and might require a large amount of hyperparameter tuning.

% Our approach is in particular based on the One Class SVM proposed by Scholk{\"o}pf et al. which estimates a user
% defined quantile level set for a collection of training data by solving a convex quadratic program  to find support
% vectors. The method has been theoretically shown to approximate the quantile levelset of a density estimate
% asymptotically for correctly chosen bandwidth settings and in the case of a normalized Gaussian kernel function \cite{vert2006consistency}. 
% In \cite{liu2014unsupervised}, the One Class SVM has furthermore been used for outlier detection of high-dimensional 
% image data.

\section{Problem Statement}
The goal of this work is to 1)learn a policy, such that for states visited after completion of training the difference in controls taken by the robot is as small as possible to that of the supervisor 2)the supervisor is only requested for a query in states, which the robot is mis-predicting the supervisor's control.

\noindent\textbf{Modeling Choices and Assumptions} We model the system dynamics as markovian and stochastic. The system dynamics are model to be stationary (i.e. given a state and a control the probability of the next state doesn't change over time). We model the initial state as sampled from a distribution over the state space.

We assume a known state space and set of controls. We assume access to a robot or simulator, such that we can start in some initial state and apply a series of controls and observe the following states that occur.  Lastly, we assume access to a supervisor that can provide a series of demonstrated trajectories and can be given a state visited by the robot or simulator and provide a control signal of what the supervisor would do.

\noindent\textbf{Policies and State Densities.}
We denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of \mlnote{We should discuss if we are going to consider images states or not}
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We assume Markovian dynamics, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

@@

Given a candidate policy $\pi_{\theta}$, we then use the surrogate loss function to approximately measure how "close" the policy's
returned control input $\pi_{\theta}(\bx)\in \mathcal{U}$ at a given state $\bx\in \mathcal{X}$ is to the supervisor's policy's control output
$\tilde{\pi}(\bx)\in \mathcal{U}$. The goal is to produce a that minimizes the surrgoate loss between relative to the supervisor's policy.

@@

 \vspace{-2ex}
\begin{align}\label{eq:LFD_obj}
\underset{\theta}{\min} \: E_{p(\bx|\theta)} [l(\pi_\theta(\bx),\tilde{\pi}(\bx))].
\end{align}
with $p(\bx|\theta)$ from Eq. \ref{eq:density}.

>>>>

\author{Michael Laskey$^1$, Sam Staszak$^3$, Wesley Yu-Shu Hsieh$^1$, Jeffrey Mahler$^1$, \\ Florian T. Pokorny$^1$, Anca D. Dragan$^1$, and Ken Goldberg$^{1,2}$% <-this % stops a space
\thanks{$^1$ Department of Electrical Engineering and Computer Sciences; {\small \{mdlaskey,iamwesleyhsieh, ftpokorny, anca\}@berkeley.edu}}%
\thanks{$^2$ Department of Industrial Engineering and Operations Research; {\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley;  Berkeley, CA 94720, USA}%
}
\begin{document}

@@

\begin{abstract}
Online algorithms for robot model-free robot learning from demonstration (such as DAgger) can learn policies for problems where the system
dynamics and the cost function are unknown. However, during learning they impose a burden on supervisors to provide labels (control
signals) as the robot encounters new states when iteratively executing its current best policy. We introduce the SHIV
algorithm (Svm-based reduction in Human InterVention), which reduces this burden by enabling the robot to \emph{only request supervision for risky states}. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to non-stationary high dimensional state distributions.  To facilitate scaling and outlier rejection, risk is based on distance to an approximate level set boundary defined by a One Class support vector machine.  We compare SHIV with DAgger in three policy learning experiments: 1) a driving simulator with a 27936D visual HOG feature state representation, 2) 4DOF robot arm push-grasping in clutter (simulation with Box2D) with a 22D state representation  of joint angles and the pose of each object, and 3) surgical needle insertion with a 16D state represented spatial transform.  Results suggest that SHIV can reduce the number of queries up to 70$\%$\adnote{double check this}.

@@

In model-free robot learning from demonstration, a robot learns to perform a task, such as driving or grasping an object in a cluttered environment (Fig. \ref{fig:teaser}), from examples provided by a  supervisor, usually a human.  In such problems, the robot does not have access to neither the cost function that it should optimize, nor the dynamics model. The former happens when it is difficult for a human to specify how to trade-off various aspects that matter, like when the robot is reaching for a goal object while not pushing objects off a table(Fig. \ref{fig:teaser}(b).  The latter happens when either the system or the interaction with the world is difficult to characterize with the given observations, like in Fig. \ref{fig:teaser}(a) the robot must stay on the track but only has access to visual information from a bird's eye view.

Rather than learning the cost function and the dynamics model, and then using optimization to produce a policy for the task, the robot learns the policy directly from supervisor examples mapping states to controls~\cite{argall2009survey}.

In the offline learning from demonstration setting, the robot learns the policy based on a batch of examples, and then executes it to achieve the task.  During execution, a small error can accumulate and push the robot away from the region of the state space where it has seen examples. The learned policy might not be helpful in unfamiliar states, leading to unrecoverable failures.  For example, a robot driving may be trained on examples driving safely down the center of a lane, but even slight deviations will put the robot into states near the side of the road where its policy could  fail~\cite{pomerleau1989alvinn}. The number of errors made by the robot, in the worst case,  can scale quadratically with the time horizon of the task~\cite{ross2010efficient}.

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth]{figures/teaser.pdf}
\caption{ 
SHIV  reduces the number of queries to the supervisor in our three test domains: a) a driving simulator where the goal is to learn a controller on HOG features extracted from synthetic images to keep a car on a polygonal track; b) push-grasping in clutter in a physics simulator, where the goal is to learn a controller from human demonstrations to grasp the green object while not letting the red square objects fall off the gray boundary representing a table; c) learning a controller for the Da Vinci Research Kit from a master surgeon's demonstrations of correcting a needle before suturing in surgery; d) cost of SHIV and DAgger's policy on the grasping in clutter example versus states labeled.
}
\vspace*{-10pt}
\label{fig:teaser}
\end{figure}

Online learning from demonstration approaches such as DAgger address this issue by iteratively gathering more examples from the supervisor in states the robot is likely to visit~\cite{grollman2007dogged,ross2010efficient,ross2010reduction}. DAgger learns a series of policies. At each iteration, the robot trains a policy based on the existing examples, then rolls out (executes) that policy and the supervisor provides demonstrations for all states the robot visits. The new examples are aggregated with the old examples for the next iteration. DAgger and related algorithms have been applied in a wide range of applications, from quadrotor flight to natural language to Atari games~\cite{NIPS2014_5421,duvallet2013imitation,ross2013learning}. In DAgger, the number of errors scales only linearly with the time horizon of the task.

To achieve better performance guarantees, online algorithms can significantly increase the burden on the supervisor, who now labels all states that the robot visits during training.  Our goal is to reduce supervisor burden without degrading robot performance.

Our key insight, to reduce burden, is that the robot only needs to request demonstrations for risky states. Risk arises because:  (1) states are different than the states previously encountered and used to train the current policy, i.e., they can be considered novel or outliers~\cite{hodge2004survey}; or (2) states are in a region that has been misclassified, in previous training data.
%\adnote{Ken's version below does not seem to represent what we actually do.}
%the action associated with a state under the current policy yield a resulting state that is inconsistent with the expected next state. [how do we measure 2? - Ken]

We contribute an algorithm for online learning from demonstration that can actively decide whether it is necessary to ask the supervisor to label the new states that the robot encounters. Our algorithm, SHIV  (Svm-based reduction in Human InterVention), reduces supervisor burden by only requesting supervision for risky states. SHIV uses stream-based active learning with a query selection method that evaluates risk in a manner tailored to the non-stationary state distributions the robot encounters as it iteratively executes learned policies. We build on the One Class SVM method for approximating quantile level sets \cite{scholkopf2001estimating} to provide a classification method for deciding if a state is risky or safe.

\mlnote{Ken didn't want to repeat the abstract here, so we removed this section}
\adnote{we need results here; both the highlights of the analysis we do (so that people don;t think we are blindly comparing it to dagger without asking if all these changes are necessary), as well as the main point, that shiv is better for the same budget, and refer to fig 1}

@@

\noindent\textbf{Active Learning from Demonstration}
Online learning from demonstration can be formulated as a stream-based active learning problem ~\cite{atlas1990training,cohn1994improving}. In stream based active learning, the decision of whether to query the supervisor (for a label or control signal) or not is not over the entire state space (like in traditional pool-based active learning), but on states drawn one at a time from some data source. In  online learning  from demonstration, the data source is the states encountered during learning.

@@

The  estimator is a function of the current distribution. Therefore, it no longer provides a good measure of confidence when that distribution changes, as it does when the robot is rolling out a learned policy and starts encountering further away states. Instead of relying on the estimator, our measure of risk explicitly identifies when states are drawn from a different distribution, i.e. when they are novel.

%%In SHIV, we use a query selection method that allows for a non-stationary distribution, evaluating risk based on novelty (is the state an outlier to the training distribution), as in Fig. \ref{fig:support_example}(c). Additionally, we also assign higher risk to  historically mislabeled regions, as in Fig. \ref{fig:support_example}(d). We then use a classification method based on the One Class SVM \cite{scholkopf2001estimating} to predict if a state is risky or not risky, which we we define as safe.%Our approach builds on the One Class SVM method for approximating quantile level sets \cite{}, which has already been used in novelty detection on high-dimensional image data \cite{}.

\noindent\textbf{Risk via Novelty Detection.}
One part of our risk definition is the notion that a trained model will be able to generalize within the distribution it is
trained on \cite{tokdar2010importance}, making states outside of this distribution risky. Novelty detection \cite{hodge2004survey} is concerned with recognizing when this happens: recognizing that a sample is outside of this distribution.

One approach to novelty detection is to directly estimate the underlying probability density function from which the data is sampled. If the probability of that point is low with respect to the underlying density, it will be marked as novel. However, the amount of data needed to accurately do so scales exponentially in the number of dimensions \cite{nadaraya1964estimating}.

An alternative to evaluating the probability of a data point is to measure distance to its nearest neighbors \cite{knox1998algorithms}. However, this approach was shown to be
susceptible to issues, since nearest neighbors  incorporates only local information about the data. For example, a group of outliers can be close together, but significantly far from the majority of the data and nearest neighbors would mark them as not novel \cite{hodge2004survey}.

Our approach is based on the One Class SVM proposed by Scholk{\"o}pf et al., which estimates a particular quantile level set for the training data by solving a convex quadratic program  to the find support
vectors~\cite{scholkopf2001estimating}. The method has been theoretically shown to approximate the quantile levelset of a density estimate
asymptotically for correctly chosen bandwidth settings and in the case of a normalized Gaussian kernel function \cite{vert2006consistency}. 
In \cite{liu2014unsupervised}, the One Class SVM has furthermore been used for novelty detection in high-dimensional image data. To our knowledge the use of One Class SVM to online learning from demonstration has never been done before. \mlnote{Ken, wanted to our knowledge in the last part} \adnote{you can say this to me, but not really write it in a paper; I think that perhaps you want to say, and in a new paragraph so that it stands out, "Our contribution is the addition of stream-based active learning to online learning from demonstration. To do so, we use a query selection method based on the One Class SVM.}

\begin{figure*}[t]
\centering

@@

\caption{A comparison of different query selection strategies for active learning on a non-stationary distribution. The shaded and empty circles are training data from two classes, $1$ and $0$ respectively. The red empty circles are samples from a new distribution (produced by executing the learned policy), and belong to class $0$. Typical strategies classify states close the decision boundary as risky(a), or for which a set of estimators disagree (b). Neither of these apply to our new samples in red. In contrast, we use a strategy that is amenable to non-stationary distributions by classifying  novel states as safe (i.e not risky) (c) and states in historically mislabeled regions. }
\label{fig:support_example}
\end{figure*}

\section{Problem Statement}
The goal of this work is to learn a policy that matches that of the supervisor's while asking the supervisor for as few examples as possible.

\mlnote{Ken is a big fan of  discussing the assumptions and model choices without notation before everything else, so a reader can quickly determine if this method suits there problem}

\noindent\textbf{Modeling Choices and Assumptions}  We model the system dynamics as markovian, stochastic, and stationary. Stationary meaning  given a state and a control the probability of the next state doesn't change over time. We model the initial state as sampled from a distribution over the state space.

We assume a known state space and set of controls. We assume access to a robot or simulator, such that we  can sample from the state sequences induced by a sequence of controls.   Lastly, we assume access to a supervisor who can, given a state, provide the desired control signal label.

\noindent\textbf{Policies and State Densities.}
We denote by $\mathcal{X}$ the set consisting of observable states for a robot task, consisting, for example, of \mlnote{We should discuss if we are going to consider images states or not}
high-dimensional vectors corresponding to images from a camera, or robot joint angles and object poses in the environment.
We furthermore consider a set $\mathcal{U}$ of allowable control inputs for the robot, which can be discrete or
continuous. We model dynamics, as Markovian, such that the probability of state $\mathbf{x_{t+1}}\in
\mathcal{X}$ can be determined from the previous state $\mathbf{x}_t\in\mathcal{X}$ and control input $\mathbf{u}_t\in
\mathcal{U}$: 
$$p(\bx_{t+1}|\bu_{t},\bx_{t}, \ldots, \bu_{0}, \bx_{0})=p(\bx_{t+1}|\bu_{t}, \bx_t)$$
We assume a probability density over initial states $p(\bx_0)$.
%We denote the probability density over the initial state also by $p:\mathcal{X}\to \mathbb{R}$.

@@

Given a candidate policy $\pi_{\theta}$, we then use the surrogate loss function to approximately measure how "close" the policy's
returned control input $\pi_{\theta}(\bx)\in \mathcal{U}$ at a given state $\bx\in \mathcal{X}$ is to the supervisor's policy's control output
$\tilde{\pi}(\bx)\in \mathcal{U}$. The goal is to produce a policy that minimizes the surrgoate loss between relative to the supervisor's policy.

@@

 \vspace{-2ex}
\begin{align}\label{eq:LFD_obj}
\underset{\theta}{\min} \: E_{p(\bx|\theta)} [l(\pi_\theta(\bx),\tilde{\pi}(\bx))]
\end{align}
with $p(\bx|\theta)$ from Eq. \ref{eq:density}.

 If the robot could learn the policy  perfectly, this state density would match the one encountered in the user examples. But if the robot makes an error, that error changes the distribution of states that the robot will visit, which can lead to states that are far away from any examples and difficult to generalize to \cite{pomerleau1989alvinn}.Thus, motivating iterative algorithms like DAgger and now SHIV, which iterate between learning a policy and then the supervisor providing feedback. The feedback is in the form of control signals on states sampled from the robot's new distribution of states.
